---
title: "Assessing The Impact of Audience of NBA Player Performance"
output: html_notebook
---

Libraries
```{r include=FALSE}
library(tidyverse)
library(ggplot2)
library(ggrepel)
library(tidyr)
library(stats)
library(scales)
```

# Load Data

Set working directory
```{r}
setwd('C:/Users/alexs/OneDrive - University of Calgary/School/1. DATA 501')
```

Read CSV
```{r}
player_game_data <- read.csv("data/player_game_data.csv", header=T, encoding="UTF-8")
team_game_data <- read.csv("data/team_game_data.csv", header=T, encoding="UTF-8")
```

# EDA

Explore audience distributions
```{r}
ggplot(data=team_game_data, aes(x=as.numeric(attendance))) +
  geom_histogram(binwidth = 500) + 
  facet_wrap(~season)
```

Explore Home vs Away Performance - Boxplots (No Bubble)
```{r}
ggplot(data = filter(team_game_data, season==1 & BubbleGame==0), aes(x = PTS, y = HomeAway)) +
  geom_boxplot() + 
  facet_wrap(~Team)

ggplot(data = filter(team_game_data, season==2 & BubbleGame==0), aes(x = PTS, y = HomeAway)) +
  geom_boxplot() + 
  facet_wrap(~Team)

ggplot(data = filter(team_game_data, season==3 & BubbleGame==0), aes(x = PTS, y = HomeAway)) +
  geom_boxplot() + 
  facet_wrap(~Team)

ggplot(data = filter(team_game_data, season==4 & BubbleGame==0), aes(x = PTS, y = HomeAway)) +
  geom_boxplot() + 
  facet_wrap(~Team)
```

Explore Home vs Away Performance - T-Tests (No Bubble)
```{r}
team_season_df <- team_game_data %>%
  distinct(team_season) %>%
  select(team_season)

ttest_df <- data.frame()

for (i in 1:nrow(team_season_df)){
  ttest_i <- t.test(PTS ~ HomeAway, data=filter(team_game_data, team_season==team_season_df[i,] & BubbleGame==0))
  ttest_i <- data.frame(pvalue = ttest_i$p.value,
                        Away = ttest_i$estimate[[1]],
                        Home = ttest_i$estimate[[2]])
  ttest_df <- rbind(ttest_df, ttest_i)
}

ttest_df <- cbind(team_season_df, ttest_df)

rm(ttest_i)
```

```{r}
ttest_df$Advantage[ttest_df$pvalue<=0.10 & ttest_df$Home>ttest_df$Away]="Significant Home Advantage"
ttest_df$Advantage[ttest_df$pvalue>0.10 & ttest_df$Home>ttest_df$Away]="Non-significant Home Advantage"
ttest_df$Advantage[ttest_df$pvalue<=0.10 & ttest_df$Home<ttest_df$Away]="Significant Away Advantage"
ttest_df$Advantage[ttest_df$pvalue>0.10 & ttest_df$Home<ttest_df$Away]="Non-significant Away Advantage"
```

```{r}
ttest_df <- merge(x = ttest_df, 
                  y = team_game_data %>%
                        group_by(team_season) %>%
                        summarise(attendance = mean(attendance)),
                  by.x = "team_season",
                  by.y = "team_season")
```

```{r}
table(ttest_df$Advantage)
```

```{r}
table(ttest_df$Advantage[ttest_df$attendance<10000])
table(ttest_df$Advantage[ttest_df$attendance>10000])
```

```{r}
filter(ttest_df, Advantage=="Non-significant Away Advantage" & attendance > 10000)
```

Explore Home vs Away Performance - Bubble
```{r}
games_played <- team_game_data %>%
  filter(BubbleGame==1) %>%
  group_by(Team) %>%
  summarise(Games = sum(BubbleGame))
```

```{r}
ggplot(data = filter(team_game_data, BubbleGame==1), aes(x = PTS, y = HomeAway)) +
  geom_boxplot() + 
  geom_text(data=games_played, 
            mapping=aes(x=80, y=1, label=Games),
            inherit.aes = F) + 
  facet_wrap(. ~ Team)
```

```{r}
bubble_ttest_df <- data.frame()

for (i in 1:nrow(games_played)){
  bubble_ttest_i <- t.test(PTS ~ HomeAway, data=filter(team_game_data, BubbleGame==1 & Team==as.character(games_played[i,1])))
  bubble_ttest_i <- data.frame(pvalue = bubble_ttest_i$p.value,
                        Away = bubble_ttest_i$estimate[[1]],
                        Home = bubble_ttest_i$estimate[[2]])
  bubble_ttest_df <- rbind(bubble_ttest_df, bubble_ttest_i)
}

bubble_ttest_df <- cbind(games_played, bubble_ttest_df)

rm(bubble_ttest_i)
```

```{r}
bubble_ttest_df$Advantage[bubble_ttest_df$pvalue<=0.10 & bubble_ttest_df$Home>bubble_ttest_df$Away]="Significant Home Advantage"
bubble_ttest_df$Advantage[bubble_ttest_df$pvalue>0.10 & bubble_ttest_df$Home>bubble_ttest_df$Away]="Non-significant Home Advantage"
bubble_ttest_df$Advantage[bubble_ttest_df$pvalue<=0.10 & bubble_ttest_df$Home<bubble_ttest_df$Away]="Significant Away Advantage"
bubble_ttest_df$Advantage[bubble_ttest_df$pvalue>0.10 & bubble_ttest_df$Home<bubble_ttest_df$Away]="Non-significant Away Advantage"
```

```{r}
table(bubble_ttest_df$Advantage)
```

## Proportion Tests for Home Court Advantage

All Games
```{r}
ha <- data.frame(xtabs(~HomeAway + Team, data=team_game_data))
wl_ha <- data.frame(xtabs(~WinLoss + HomeAway + Team, data=team_game_data))

team_prop <- left_join(wl_ha,
                       ha,
                       by = c("HomeAway", "Team"))

team_prop <- team_prop %>%
  filter(WinLoss=="Win")

team_prop_away <- team_prop %>%
  filter(HomeAway=="Away")

team_prop_home <- team_prop %>%
  filter(HomeAway=="Home") 

rm(wl_ha, ha)

prop_test_df <- data.frame()

for (i in 1:30){
  
  prop_test_i <- prop.test(c(team_prop_home[i,4], team_prop_away[i,4]), c(team_prop_home[i,5], team_prop_away[i,5]))
  prop_test_i <- data.frame(pvalue = prop_test_i$p.value,
                        Home = prop_test_i$estimate[[1]],
                        Away = prop_test_i$estimate[[2]])
  prop_test_df <- rbind(prop_test_df, prop_test_i)
}

prop_test_df <- cbind(team_prop_home[,c(3:5)], team_prop_away[,c(4:5)], prop_test_df)

rm(prop_test_i)

prop_test_df
prop.test(c(sum(team_prop_home[,4]), sum(team_prop_away[,4])), c(sum(team_prop_home[,5]), sum(team_prop_away[,5])))
```

All empty arena games
```{r}
ha <- data.frame(xtabs(~HomeAway + Team, data=filter(team_game_data, empty==1)))
wl_ha <- data.frame(xtabs(~WinLoss + HomeAway + Team, data=filter(team_game_data, empty==1)))

team_prop <- left_join(wl_ha,
                       ha,
                       by = c("HomeAway", "Team"))

team_prop <- team_prop %>%
  filter(WinLoss=="Win")

team_prop_away <- team_prop %>%
  filter(HomeAway=="Away")

team_prop_home <- team_prop %>%
  filter(HomeAway=="Home") 

rm(wl_ha, ha)

prop_test_df <- data.frame()

for (i in 1:30){
  
  prop_test_i <- prop.test(c(team_prop_home[i,4], team_prop_away[i,4]), c(team_prop_home[i,5], team_prop_away[i,5]))
  prop_test_i <- data.frame(pvalue = prop_test_i$p.value,
                        Home = prop_test_i$estimate[[1]],
                        Away = prop_test_i$estimate[[2]])
  prop_test_df <- rbind(prop_test_df, prop_test_i)
}

prop_test_df <- cbind(team_prop_home[,c(3:5)], team_prop_away[,c(4:5)], prop_test_df)

rm(prop_test_i)

prop_test_df
prop.test(c(sum(team_prop_home[,4]), sum(team_prop_away[,4])), c(sum(team_prop_home[,5]), sum(team_prop_away[,5])))
```

All non-empty games
```{r}
ha <- data.frame(xtabs(~HomeAway + Team, data=filter(team_game_data, empty==0)))
wl_ha <- data.frame(xtabs(~WinLoss + HomeAway + Team, data=filter(team_game_data, empty==0)))

team_prop <- left_join(wl_ha,
                       ha,
                       by = c("HomeAway", "Team"))

team_prop <- team_prop %>%
  filter(WinLoss=="Win")

team_prop_away <- team_prop %>%
  filter(HomeAway=="Away")

team_prop_home <- team_prop %>%
  filter(HomeAway=="Home") 

rm(wl_ha, ha)

prop_test_df <- data.frame()

for (i in 1:30){
  
  prop_test_i <- prop.test(c(team_prop_home[i,4], team_prop_away[i,4]), c(team_prop_home[i,5], team_prop_away[i,5]))
  prop_test_i <- data.frame(pvalue = prop_test_i$p.value,
                        Home = prop_test_i$estimate[[1]],
                        Away = prop_test_i$estimate[[2]])
  prop_test_df <- rbind(prop_test_df, prop_test_i)
}

prop_test_df <- cbind(team_prop_home[,c(3:5)], team_prop_away[,c(4:5)], prop_test_df)

rm(prop_test_i)

prop_test_df
prop.test(c(sum(team_prop_home[,4]), sum(team_prop_away[,4])), c(sum(team_prop_home[,5]), sum(team_prop_away[,5])))
```

Bubble Games
```{r}
ha <- data.frame(xtabs(~HomeAway + Team, data=filter(team_game_data, BubbleGame==1)))
wl_ha <- data.frame(xtabs(~WinLoss + HomeAway + Team, data=filter(team_game_data, BubbleGame==1)))

team_prop <- left_join(wl_ha,
                       ha,
                       by = c("HomeAway", "Team"))

team_prop <- team_prop %>%
  filter(WinLoss=="Win")

team_prop_away <- team_prop %>%
  filter(HomeAway=="Away")

team_prop_home <- team_prop %>%
  filter(HomeAway=="Home") 

rm(wl_ha, ha)

prop_test_df <- data.frame()

for (i in 1:22){
  
  prop_test_i <- prop.test(c(team_prop_home[i,4], team_prop_away[i,4]), c(team_prop_home[i,5], team_prop_away[i,5]))
  prop_test_i <- data.frame(pvalue = prop_test_i$p.value,
                        Home = prop_test_i$estimate[[1]],
                        Away = prop_test_i$estimate[[2]])
  prop_test_df <- rbind(prop_test_df, prop_test_i)
}

prop_test_df <- cbind(team_prop_home[,c(3:5)], team_prop_away[,c(4:5)], prop_test_df)

rm(prop_test_i)

prop_test_df
prop.test(c(sum(team_prop_home[,4]), sum(team_prop_away[,4])), c(sum(team_prop_home[,5]), sum(team_prop_away[,5])))
```

Non-bubble playoff games
```{r}
ha <- data.frame(xtabs(~HomeAway + Team, data=filter(team_game_data, Playoff==1)))
wl_ha <- data.frame(xtabs(~WinLoss + HomeAway + Team, data=filter(team_game_data, Playoff==1)))

team_prop <- left_join(wl_ha,
                       ha,
                       by = c("HomeAway", "Team"))

team_prop <- team_prop %>%
  filter(WinLoss=="Win") 

team_prop_away <- team_prop %>%
  filter(HomeAway=="Away")

team_prop_home <- team_prop %>%
  filter(HomeAway=="Home") 

rm(wl_ha, ha)

prop_test_df <- data.frame()

# for (i in 1:30){
#   
#   prop_test_i <- prop.test(c(team_prop_home[i,4], team_prop_away[i,4]), c(team_prop_home[i,5], team_prop_away[i,5]))
#   prop_test_i <- data.frame(pvalue = prop_test_i$p.value,
#                         Home = prop_test_i$estimate[[1]],
#                         Away = prop_test_i$estimate[[2]])
#   prop_test_df <- rbind(prop_test_df, prop_test_i)
# }
# 
# prop_test_df <- cbind(team_prop_home[,c(3:5)], team_prop_away[,c(4:5)], prop_test_df)
# 
# rm(prop_test_i)
# 
# prop_test_df
prop.test(c(sum(team_prop_home[,4]), sum(team_prop_away[,4])), c(sum(team_prop_home[,5]), sum(team_prop_away[,5])))
```

Empty, non-bubble games
```{r}
ha <- data.frame(xtabs(~HomeAway + Team, data=filter(team_game_data, BubbleGame==0 & empty==1)))
wl_ha <- data.frame(xtabs(~WinLoss + HomeAway + Team, data=filter(team_game_data, BubbleGame==0 & empty==1)))

team_prop <- left_join(wl_ha,
                       ha,
                       by = c("HomeAway", "Team"))

team_prop <- team_prop %>%
  filter(WinLoss=="Win")

team_prop_away <- team_prop %>%
  filter(HomeAway=="Away" & !(Team %in% c("Houston Rockets", "New Orleans Pelicans")))

team_prop_home <- team_prop %>%
  filter(HomeAway=="Home" & !(Team %in% c("Houston Rockets", "New Orleans Pelicans"))) 

rm(wl_ha, ha)

prop_test_df <- data.frame()

for (i in 1:28){
  
  prop_test_i <- prop.test(c(team_prop_home[i,4], team_prop_away[i,4]), c(team_prop_home[i,5], team_prop_away[i,5]))
  prop_test_i <- data.frame(pvalue = prop_test_i$p.value,
                        Home = prop_test_i$estimate[[1]],
                        Away = prop_test_i$estimate[[2]])
  prop_test_df <- rbind(prop_test_df, prop_test_i)
}

prop_test_df <- cbind(team_prop_home[,c(3:5)], team_prop_away[,c(4:5)], prop_test_df)

rm(prop_test_i)

prop_test_df
prop.test(c(sum(team_prop_home[,4]), sum(team_prop_away[,4])), c(sum(team_prop_home[,5]), sum(team_prop_away[,5])))
```

The higher home vs away is likely due to the nature of playoffs - higher seeded
teams play more games at home. In the bubble, better teams play more games
classified as "home".

Overall, there is definitely a statistically significant home court advantage.
This advantage is most prevalent in arenas with fans. There is still a home court 
advantage in empty arenas outside the bubble, likely due to the effects travel 
has on teams, as well as familiarity to the environment they play in. While fans
play an important role in the home court advantage, it seems that the advantage 
prevails, albeit less strongly, even in empty arenas. At an individual team level
there are no significant advantages in empty arenas, however there are mostly 
observational home advantages. In non-empty arenas, all teams have a significant
home court advantage

# Player EDA

## Most empty games played
```{r}
games_empty <- data.frame(xtabs(~Starters + empty, data=filter(player_game_data, empty==1)))
```


## Nikola Jokic
```{r}
nikola <- filter(player_game_data, Starters=="Nikola Jokic")
xtabs(~ season + empty, data=nikola)

hist(filter(nikola, season==2 | season==3)$empty)
hist(nikola$BPM)

ggplot(data=filter(nikola, season==2 | season==3), aes(x=empty)) +
  geom_histogram(bins=3) +
  xlab("Empty (0 is Not Empty, 1 is Empty)") +
  ggtitle("Games Played in Empty vs Non-Empty Environments")
```

Examine trends
```{r}
bpm_ts <- nikola %>%
  select(BPM, FT., PTS, MP)

bpm_ts <- ts(bpm_ts)

plot(bpm_ts[,1])
acf(bpm_ts[,1])
```

Test performance with season 2 and 3 data
```{r}
t.test(BPM~empty, data=filter(nikola, season==2 | season==3))
t.test(FT.~empty, data=filter(nikola, season==2 | season==3))
t.test(PTS~empty, data=filter(nikola, season==2 | season==3))
t.test(MP~empty, data=filter(nikola, season==2 | season==3))
```

## Turn the above into a function where you can enter a player and see results

```{r}
player_performance <- function(player){
  
  # create data frame of selected player
  player_df <- filter(player_game_data, Starters==player)
  
  # explore distribution of empty arena games and BPM
  empty_dist <- hist(filter(player_df, season==2 | season==3)$empty)
  BPM_dist <- hist(player_df$BPM)
  
  # analyze BPM trends
  bpm_ts <- player_df %>%
    select(BPM) %>%
    filter(BPM != "NA")
  bpm_ts <- ts(bpm_ts)
  ts_plot <- bpm_ts
  acf_plot <- acf(bpm_ts)
  
  # perform t-tests
  BPM_test <- t.test(BPM~empty, data=filter(player_df, season==2 | season==3))
  FT._test <- t.test(FT.~empty, data=filter(player_df, season==2 | season==3))
  PTS_test <- t.test(PTS~empty, data=filter(player_df, season==2 | season==3))
  MP_test <- t.test(MP~empty, data=filter(player_df, season==2 | season==3))
  
  result <- list(empty_dist, 
                 BPM_dist,
                 ts_plot,
                 acf_plot,
                 BPM_test,
                 FT._test,
                 PTS_test,
                 MP_test)
  
  return(result)
}
```

```{r}
lebron <- player_performance("LeBron James")
plot(lebron[[3]])
lebron[5:8]
```

# How can we use data science techniques to determine the impact that the fans in attendance have on NBA players?

Create Function
```{r}
player_impact <- function(player_name, data){
  data <- data %>% 
    filter(Starters==player_name & season %in% c(2,3)) %>%
    select(Starters, PTS, BPM, FT, FTA, empty)
  
  t_test <- t.test(BPM ~ empty, data = data)
  
  t_test_pts <- t.test(PTS ~ empty, data = data)
  
  prop_test <- prop.test(c(sum(filter(data, empty==0)$FT, na.rm=T), sum(filter(data, empty==1)$FT, na.rm=T)), 
                        c(sum(filter(data, empty==0)$FTA, na.rm=T), sum(filter(data, empty==1)$FTA, na.rm=T)))
  
  return_data <- data.frame(Player = player_name,
                            BPM_nonempty = t_test$estimate[[1]],
                            BPM_empty = t_test$estimate[[2]],
                            p_value = t_test$p.value,
                            PTS_nonempty = t_test_pts$estimate[[1]],
                            PTS_empty = t_test_pts$estimate[[2]],
                            p_value1 = t_test_pts$p.value,
                            FT._nonempty = prop_test$estimate[[1]],
                            FT._empty = prop_test$estimate[[2]],
                            p_value2 = prop_test$p.value)
  
  return(return_data)
}
```

Loop Through Each Player
```{r include=FALSE}
outliers <- boxplot(player_game_data$BPM)$out

filter(player_game_data, !(BPM %in% outliers))

player_list <- player_game_data %>%
      filter(empty==1 & season %in% c(2,3), MP >= 10) %>%
      select(Starters) %>%
      count(Starters) %>%
      filter(n>=25) %>%
      pull(Starters)

player_list2 <- player_game_data %>%
      filter(empty==0 & season %in% c(2,3), MP >= 10) %>%
      select(Starters) %>%
      count(Starters) %>%
      filter(n>=25) %>%
      pull(Starters)

player_list <- as.character(intersect(player_list, player_list2))

player_impact_df <- player_impact(player_list[1], player_game_data)

for (i in c(1:238)){
  i_data <- player_impact(player_list[i], player_game_data)
  player_impact_df <- rbind(player_impact_df, i_data)
}
```

Flag Significant differences
```{r}
player_impact_df$SigBPM[player_impact_df$p_value<=0.1 & player_impact_df$BPM_nonempty < player_impact_df$BPM_empty] = "Significantly Increased BPM"
player_impact_df$SigBPM[player_impact_df$p_value>=0.1& player_impact_df$BPM_nonempty < player_impact_df$BPM_empty] = "Insignificantly Increased BPM"
player_impact_df$SigBPM[player_impact_df$p_value<=0.1 & player_impact_df$BPM_nonempty > player_impact_df$BPM_empty] = "Significantly Decreased BPM"
player_impact_df$SigBPM[player_impact_df$p_value>=0.1& player_impact_df$BPM_nonempty > player_impact_df$BPM_empty] = "Insignificantly Decreased BPM"
player_impact_df$SigBPM[player_impact_df$p_value>=0.9] = "Nearly Equal BPM"

player_impact_df$SigPTS[player_impact_df$p_value1<=0.1 & player_impact_df$PTS_nonempty < player_impact_df$PTS_empty] = "Significantly Increased PTS"
player_impact_df$SigPTS[player_impact_df$p_value1>=0.1& player_impact_df$PTS_nonempty < player_impact_df$PTS_empty] = "Insignificantly Increased PTS"
player_impact_df$SigPTS[player_impact_df$p_value1<=0.1 & player_impact_df$PTS_nonempty > player_impact_df$PTS_empty] = "Significantly Decreased PTS"
player_impact_df$SigPTS[player_impact_df$p_value1>=0.1& player_impact_df$PTS_nonempty > player_impact_df$PTS_empty] = "Insignificantly Decreased PTS"
player_impact_df$SigPTS[player_impact_df$p_value1>=0.9] = "Nearly Equal PTS"

player_impact_df$SigFT.[player_impact_df$p_value2<=0.1 & player_impact_df$FT._nonempty < player_impact_df$FT._empty] = "Significantly Increased FT%"
player_impact_df$SigFT.[player_impact_df$p_value2>=0.1 & player_impact_df$FT._nonempty < player_impact_df$FT._empty] = "Insignificantly Increased FT%"
player_impact_df$SigFT.[player_impact_df$p_value2<=0.1 & player_impact_df$FT._nonempty > player_impact_df$FT._empty] = "Significantly Decreased FT%"
player_impact_df$SigFT.[player_impact_df$p_value2>=0.1 & player_impact_df$FT._nonempty > player_impact_df$FT._empty] = "Insignificantly Decreased FT%"
player_impact_df$SigFT.[player_impact_df$p_value2>=0.9] = "Nearly Equal FT%"
```

```{r}
bpm_table <- table(player_impact_df$SigBPM)
pts_table <- table(player_impact_df$SigPTS)
ft._table <- table(player_impact_df$SigFT.)

total_table <- data.frame(rbind(bpm_table, pts_table, ft._table))
```

Find Percentages of total
```{r}
bpm_vector <- c(bpm_table[[1]],
  bpm_table[[2]],
  bpm_table[[3]],
  bpm_table[[4]],
  bpm_table[[5]])

pts_vector <- c(pts_table[[1]],
  pts_table[[2]],
  pts_table[[3]],
  pts_table[[4]],
  pts_table[[5]])

ft._vector <- c(ft._table[[1]],
  ft._table[[2]],
  ft._table[[3]],
  ft._table[[4]],
  ft._table[[5]])
```

```{r}
bpm_perc <- percent(bpm_vector[1:5]/sum(bpm_vector))
pts_perc <- percent(pts_vector[1:5]/sum(pts_vector))
ft._perc <- percent(ft._vector[1:5]/sum(ft._vector))

perc_table <- data.frame(rbind(bpm_perc, pts_perc, ft._perc))
colnames(perc_table) <- c("Nearly Equal",
                       "Insignificantly Decreased",
                       "Insignificantly Increased",
                       "Significantly Decreased",
                       "Significantly Increased")

rownames(perc_table) <- c("BPM", "PTS", "FT.")

perc_table
```

# League Level Results

Testing points in non-empty vs empty
```{r}
ggplot(filter(team_game_data, season %in% c(2, 3)), aes(x=empty)) +
  geom_bar()

# only use season 2 and 3 to even out sample size in each group
# there are still twice as many non-empty games vs empty games

t.test(PTS ~ empty, data=filter(player_game_data, season %in% c(2,3)))
t.test(PTS ~ empty, data=filter(team_game_data, season %in% c(2,3)))
```

Testing free throw percentage in non-empty vs empty
```{r}
ftm_empty <- sum(filter(team_game_data, empty==1 & season %in% c(2,3))$FT, na.rm=T)
ftm_nonempty <- sum(filter(team_game_data, empty==0 & season %in% c(2,3))$FT, na.rm=T)
fta_empty <- sum(filter(team_game_data, empty==1 & season %in% c(2,3))$FTA, na.rm=T)
fta_nonempty <- sum(filter(team_game_data, empty==0 & season %in% c(2,3))$FTA, na.rm=T)

prop.test(c(ftm_empty, ftm_nonempty), c(fta_empty, fta_nonempty))
```

Prop tests with past probabilities of winning
```{r}
xtabs(~ HomeAway + WinLoss, data = filter(team_game_data, empty==0 & Playoff==0 & season %in% c(1,2)))
xtabs(~ HomeAway + WinLoss, data = filter(team_game_data, empty==1 & Playoff==0))
xtabs(~ HomeAway + WinLoss, data = filter(team_game_data, BubbleGame==1))

# Regular season winning percentage compared to past 10 years winning percentage
prop.test(x = c(937, 1264),
          n = c(1264+937, 1264+937),
          p = c(0.4104, 0.5896),
          correct = F)

# Regular season winning percentage in empty arenas compared to past 10 years winning percentage
prop.test(x = c(290, 303),
          n = c(290+303, 290+303),
          p = c(0.4104, 0.5896),
          correct = F)

# Bubble winning percentage compared to past 10 years winning percentage
prop.test(x = c(82, 90),
          n = c(172, 172),
          p = c(0.4104, 0.5896),
          correct = F)
```

# Are certain players “big game performers”? (Playoff vs Non-Playoff Performance)

```{r}
#Select only the relevant columns for the tests
playoffs_player <- player_game_data %>% select(Starters,Playoff,BPM,Win,Season,PTS,FT.,FT,FTA,attendance)


#Create a dataframe with all unique players
unique_players <- unique(playoffs_player$Starters)
unique_players <- data.frame(unique_players)
unique_players <- unique_players %>% filter(unique_players != 'Álex Abrines' & unique_players != 'J.R. Smith' & unique_players != 'Joakim Noah' & unique_players != 'Donta Hall' & unique_players != 'Lance Thomas' & unique_players != 'Jamal Crawford' & unique_players != 'Kyle Alexander')


#Create columns for total games played, and ttal playoff games played

for (i in 1:nrow(unique_players)){
  player_filter <- playoffs_player %>% filter(Starters == unique_players$unique_players[i])
  unique_players$totgames[i] <- nrow(player_filter)
  
  player_filter <- playoffs_player %>% filter(Starters == unique_players$unique_players[i] & Playoff == 1)
  unique_players$totplayoffgames[i] <- nrow(player_filter)
  
}

#Add total games and total playoff games played to the original dataset
playoffs_player$totgames <- unique_players$totgames[cbind(match(playoffs_player$Starters,unique_players$unique_players))]

playoffs_player$totplayoffgames <- unique_players$totplayoffgames[cbind(match(playoffs_player$Starters,unique_players$unique_players))]

#Function to run tests for playoff games
playoffs <- function(player,data){

funcfilter <- data %>% filter(Starters == player)

funcfilter$BPM <- as.numeric(funcfilter$BPM)
funcfilter$PTS <- as.numeric(funcfilter$PTS)
funcfilter$FT <- as.numeric(funcfilter$FT)
funcfilter$FTA <- as.numeric(funcfilter$FTA)
funcfilter$Playofftest <- as.factor(funcfilter$Playoff)

x1 <- funcfilter %>% filter(FTA!= 0 & Playofftest == 0)
x2 <- funcfilter %>% filter(FTA!= 0 & Playofftest == 1)

boxplotbpm <- ggplot(data = funcfilter, aes(x = Playofftest, y = BPM)) + geom_boxplot()
boxplotpts <- ggplot(data = funcfilter, aes(x = Playofftest, y = PTS)) + geom_boxplot()
boxplotft <- ggplot(data = funcfilter, aes(x = Playofftest, y = FT.)) + geom_boxplot()

playoff.pts <- t.test(data = funcfilter, PTS~Playofftest, alternative = "two.sided", mu = 0)$p.value
playoff.bpm <- t.test(data = funcfilter, BPM~Playofftest, alternative = "two.sided", mu = 0)$p.value
playoff.FT <- prop.test(x = c(sum(x1$FT),sum(x2$FT)), n = c(sum(x1$FTA),sum(x2$FTA)))$p.value

estptsplayoffs <- t.test(data = funcfilter, PTS~Playofftest, alternative = "two.sided", mu = 0)$estimate
estbpm <- t.test(data = funcfilter, BPM~Playofftest, alternative = "two.sided", mu = 0)$estimate
estft <- prop.test(x = c(sum(x1$FT),sum(x2$FT)), n = c(sum(x1$FTA),sum(x2$FTA)), alternative = "two.sided")$estimate


meanpointsplayoffs <- estptsplayoffs[[2]]
meanpointsnonplayoffs <- estptsplayoffs[[1]]
meanbpmplayoffs <- estbpm[[2]]
meanbpmnonplayoffs <- estbpm[[1]]
propftplayoffs <- estft[[2]]
propftnonplayoffs <- estft[[1]]

results <- list(boxplotbpm,boxplotpts,boxplotft,playoff.pts,playoff.bpm,playoff.FT,meanpointsplayoffs,meanpointsnonplayoffs,meanbpmplayoffs,meanbpmnonplayoffs,propftplayoffs,propftnonplayoffs)

return(results)
}

#CLT condition for normality
unique_playersfilt <- unique_players %>% filter(totplayoffgames > 25 & totgames > 25)

for (i in 1:nrow(unique_playersfilt)){
  
  results <- playoffs(unique_playersfilt$unique_players[i],playoffs_player)
  
  unique_playersfilt$ptspvalue[i] <- results[4]
  unique_playersfilt$bpmpvalue[i] <- results[5]
  unique_playersfilt$ftpvalue[i] <- results[6]
  unique_playersfilt$meanptsplayoffs[i] <- results[7]
  unique_playersfilt$meanptsnonplayoffs[i] <- results[8]
  unique_playersfilt$meanbpmplayoffs[i] <- results[9]
  unique_playersfilt$meanbpmnonplayoffs[i] <- results[10]
  unique_playersfilt$propftplayoffs[i] <- results[11]
  unique_playersfilt$propftnonplayoffs[i] <- results[12]
}

as.numeric(unique_playersfilt$ptsplayoffs)
as.numeric(unique_playersfilt$bpmplayoffs)
as.numeric(unique_playersfilt$ftplayoffs)


#P Value % for points
length(unique_playersfilt$ptsplayoffs[unique_playersfilt$ptsplayoffs < 0.1]) / nrow(unique_playersfilt)
#P Value % for BPM
length(unique_playersfilt$bpmplayoffs[unique_playersfilt$bpmplayoffs < 0.1]) / nrow(unique_playersfilt)
#P Value % for FT
length(unique_playersfilt$ftplayoffs[unique_playersfilt$ftplayoffs < 0.1]) / nrow(unique_playersfilt)
```

```{r}
#Creates columns to determine significance
unique_playersfilt$PTSsignificance[unique_playersfilt$ptspvalue <= 0.1 & as.numeric(unique_playersfilt$meanptsplayoffs) > as.numeric(unique_playersfilt$meanptsnonplayoffs)] = "Significantly Increased PTS"
unique_playersfilt$PTSsignificance[unique_playersfilt$ptspvalue <= 0.1 & as.numeric(unique_playersfilt$meanptsplayoffs) < as.numeric(unique_playersfilt$meanptsnonplayoffs)] = "Significantly Decreased PTS"
unique_playersfilt$PTSsignificance[unique_playersfilt$ptspvalue > 0.1 & as.numeric(unique_playersfilt$meanptsplayoffs) > as.numeric(unique_playersfilt$meanptsnonplayoffs)] = "Insignificantly Increased PTS"
unique_playersfilt$PTSsignificance[unique_playersfilt$ptspvalue > 0.1 & as.numeric(unique_playersfilt$meanptsplayoffs) < as.numeric(unique_playersfilt$meanptsnonplayoffs)] = "Insignificantly Decreased PTS"
unique_playersfilt$PTSsignificance[unique_playersfilt$ptspvalue >= 0.9] = "Nearly Equal PTS"

unique_playersfilt$BPMsignificance[unique_playersfilt$bpmpvalue <= 0.1 & as.numeric(unique_playersfilt$meanbpmplayoffs) > as.numeric(unique_playersfilt$meanbpmnonplayoffs)] = "Significantly Increased BPM"
unique_playersfilt$BPMsignificance[unique_playersfilt$bpmpvalue <= 0.1 & as.numeric(unique_playersfilt$meanbpmplayoffs) < as.numeric(unique_playersfilt$meanbpmnonplayoffs)] = "Significantly Decreased BPM"
unique_playersfilt$BPMsignificance[unique_playersfilt$bpmpvalue > 0.1 & as.numeric(unique_playersfilt$meanbpmplayoffs) > as.numeric(unique_playersfilt$meanbpmnonplayoffs)] = "Insignificantly Increased BPM"
unique_playersfilt$BPMsignificance[unique_playersfilt$bpmpvalue > 0.1 & as.numeric(unique_playersfilt$meanbpmplayoffs) < as.numeric(unique_playersfilt$meanbpmnonplayoffs)] = "Insignificantly Decreased BPM"
unique_playersfilt$BPMsignificance[unique_playersfilt$bpmpvalue >= 0.9] = "Nearly Equal BPM"

unique_playersfilt$FTsignificance[unique_playersfilt$ftpvalue <= 0.1 & as.numeric(unique_playersfilt$propftplayoffs) > as.numeric(unique_playersfilt$propftnonplayoffs)] = "Significantly Increased FT %"
unique_playersfilt$FTsignificance[unique_playersfilt$ftpvalue <= 0.1 & as.numeric(unique_playersfilt$propftplayoffs) < as.numeric(unique_playersfilt$propftnonplayoffs)] = "Significantly Decreased FT %"
unique_playersfilt$FTsignificance[unique_playersfilt$ftpvalue > 0.1 & as.numeric(unique_playersfilt$propftplayoffs) > as.numeric(unique_playersfilt$propftnonplayoffs)] = "Insignificantly Increased FT %"
unique_playersfilt$FTsignificance[unique_playersfilt$ftpvalue > 0.1 & as.numeric(unique_playersfilt$propftplayoffs) < as.numeric(unique_playersfilt$propftnonplayoffs)] = "Insignificantly Decreased FT %"
unique_playersfilt$FTsignificance[unique_playersfilt$ftpvalue >= 0.9] = "Nearly Equal FT %"
```

```{r}
#Makes all columns from tests numeric
unique_playersfilt$ptspvalue <- as.numeric(unique_playersfilt$ptspvalue)
unique_playersfilt$bpmpvalue <- as.numeric(unique_playersfilt$bpmpvalue)
unique_playersfilt$meanptsplayoffs <- as.numeric(unique_playersfilt$meanptsplayoffs)
unique_playersfilt$meanptsnonplayoffs <- as.numeric(unique_playersfilt$meanptsnonplayoffs)
unique_playersfilt$meanbpmplayoffs <- as.numeric(unique_playersfilt$meanbpmplayoffs)
unique_playersfilt$meanbpmnonplayoffs <- as.numeric(unique_playersfilt$meanbpmnonplayoffs)
unique_playersfilt$propftplayoffs <- as.numeric(unique_playersfilt$propftplayoffs)
unique_playersfilt$propftnonplayoffs <- as.numeric(unique_playersfilt$propftnonplayoffs)

unique_playersfilt
```

Teams
```{r}
#Selects relevant columns
colnames(team_game_data)[1] <- "Team"
playoffs_team <- team_game_data %>% select(Team,Playoff,Win)

unique_team <- unique(playoffs_team$Team)
unique_team <- data.frame(unique_team)


#Counts total games and total playoff games
for (i in 1:nrow(unique_team)){
  team_filter <- playoffs_team %>% filter(Team == unique_team$unique_team[i])
  unique_team$totgames[i] <- nrow(team_filter)
  
  team_filter <- playoffs_team %>% filter(Team == unique_team$unique_team[i] & Playoff == 1)
  unique_team$totplayoffgames[i] <- nrow(team_filter)
  
}

playoffs_team$totgames <- unique_team$totgames[cbind(match(playoffs_team$Team,unique_team$unique_team))]

playoffs_team$totplayoffgames <- unique_team$totplayoffgames[cbind(match(playoffs_team$Team,unique_team$unique_team))]


#Function to run tests on team performance
playoffsteam <- function(team,data){

funcfilter <- data %>% filter(Team == team)

funcfilter$Playoff <- as.factor(funcfilter$Playoff)
funcfilter$Win <- as.numeric(funcfilter$Win)

x1 <- funcfilter %>% filter(Playoff == 0)
x2 <- funcfilter %>% filter(Playoff == 1)

playoff.Win <- prop.test(x = c(sum(x1$Win),sum(x2$Win)), n = c(nrow(x1),nrow(x2)),alternative = "two.sided")$p.value
est.Win <- prop.test(x = c(sum(x1$Win),sum(x2$Win)), n = c(nrow(x1),nrow(x2)),alternative = "two.sided")$estimate

estwinnonplayoff <- as.numeric(est.Win[[1]])
estwinplayoff <- as.numeric(est.Win[[2]])

results <- list(playoff.Win,estwinnonplayoff,estwinplayoff)
return(results)
}

for (i in 1:nrow(unique_team)){
  teamfilt <- playoffs_team %>% filter(Team == unique_team$unique_team[i])
  unique_team$totalwins[i] <- sum(teamfilt$Win)
  unique_team$winpercentage[i] <- unique_team$totalwins[i]/unique_team$totgames[i]
  
}
unique_team$condition1 <- unique_team$totplayoffgames*unique_team$winpercentage
unique_team$condition2 <- unique_team$totplayoffgames*(1-unique_team$winpercentage)

#Proportion test condition of (n*p) >= 10 and (n*(1-p)) >= 10

proportiontest_team <- unique_team %>% filter(condition1>10& condition2>10)


#Adds results as columns
for (i in 1:nrow(proportiontest_team)){
  results <- playoffsteam(proportiontest_team$unique_team[i],playoffs_team)
  
  proportiontest_team$Winpvalue[i] <- results[1]
  proportiontest_team$propwinnonplayoff[i] <- results[2]
  proportiontest_team$propwinplayoff[i] <- results[3]
}


```

```{r}
#Determines significance for each team
proportiontest_team$Winsignificance[proportiontest_team$Winpvalue <= 0.1 & as.numeric(proportiontest_team$propwinplayoff) > as.numeric(proportiontest_team$propwinnonplayoff)] = "Significantly Increased Win %"

proportiontest_team$Winsignificance[proportiontest_team$Winpvalue <= 0.1 & as.numeric(proportiontest_team$propwinplayoff) < as.numeric(proportiontest_team$propwinnonplayoff)] = "Significantly Decreased Win %"

proportiontest_team$Winsignificance[proportiontest_team$Winpvalue > 0.1 & as.numeric(proportiontest_team$propwinplayoff) > as.numeric(proportiontest_team$propwinnonplayoff)] = "Insignificantly Increased Win %"

proportiontest_team$Winsignificance[proportiontest_team$Winpvalue > 0.1 & as.numeric(proportiontest_team$propwinplayoff) < as.numeric(proportiontest_team$propwinnonplayoff)] = "Insignificantly Decreased Win %"

proportiontest_team$Winsignificance[proportiontest_team$Winpvalue >= 0.9] = "Nearly Equal Win %"
```

```{r}
proportiontest_team
```






